---
title: "Project1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Introduction
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This project makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The variables included in this dataset are:

**steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)

**date**: The date on which the measurement was taken in YYYY-MM-DD format

**interval**: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.


**Loading the required libraries for the project: dplyr and lattic**
```{r}
library("dplyr")
library("lattice")
```

*Reading the CSV in data frame*
```{r}
actdata <- read.csv("E:\\course5\\activity\\activity.csv")
```

##Below is total number of steps taken per day##
```{r}
sumstepsfordays <- actdata %>% group_by(date) %>% summarise(steps=sum(steps))
sumstepsfordays
```
##Below is histogram of total number of steps taken each day##
```{r}
hist(sumstepsfordays$steps,
     main="Histogram for Sum of steps taken per day",
     xlab="Number of Steps",
     border="blue",
     col="green", labels=TRUE, ylim=c(0,30))

```


##Below is mean and median of the total number of steps taken per day##
```{r}

normean<-mean(sumstepsfordays$steps,na.rm=TRUE)
normedian<-median(sumstepsfordays$steps,na.rm=TRUE)
```

*Mean number of steps for day is `r normean`

*Median number of steps for day is `r normedian`


```

##Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)##

```{r}
avgstepsforintervals <- actdata %>% filter(!is.na(actdata$steps)) %>% group_by(interval) %>% summarise(steps=mean(steps))
plot(x=avgstepsforintervals$interval,y=avgstepsforintervals$steps,type="l",ylim=c(0,200))

```

##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?##
```{r}
maxstepsininterval <- avgstepsforintervals %>% filter(steps==max(steps))
maxstepsininterval$interval
```

#Imputing missing values#

##Total number of missing values in the dataset##
```{r}
missingrows<-sum(!complete.cases(actdata))
missingrows

```
##Filling missing values##
```{r}
imputeddata<-
actdata %>% 
  mutate(steps= ifelse(is.na(steps), avgstepsforintervals$steps[match(actdata$interval,avgstepsforintervals$interval)], steps))


```

###As data for all days are NA, filling it with zero to make sure data trend is not changed because of imputation###
```{r}
imputeddata[as.character(imputeddata$date) == "2012-10-01", 1] <- 0


```

##Histogram of the total number of steps taken each day##
```{r}
impsumstepsfordays <- imputeddata %>% group_by(date) %>% summarise(steps=sum(steps))


hist(impsumstepsfordays$steps,
     main="Histogram for Sum of steps taken per day",
     xlab="Number of Steps",
     border="blue",
     col="green", labels=TRUE, ylim=c(0,40))

```

##Mean and median of steps taken per day in *imputed* data##
```{r}
impmean <- mean(impsumstepsfordays$steps)
impmedian <- median(impsumstepsfordays$steps)
```
*Mean of steps in imputed data is `r impmean`*

*Median of steps in imputed data is `r impmedian`*

##Below is the histogram to show the difference of data before and after imputation of missing values##
```{r}
hist(impsumstepsfordays$steps,
     main="Histogram for Sum of steps taken per day",
     xlab="Number of Steps",
     border="blue",
     col="green", labels=TRUE, ylim=c(0,40))


hist(sumstepsfordays$steps,
     main="Histogram for Sum of steps taken per day",
     xlab="Number of Steps",
     border="blue",
     col="red", labels=TRUE, ylim=c(0,30),add=T)

legend("topright", c("Imputed", "Non-imputed"), col=c("green", "red"), lwd=10)
```


##Difference between mean and median of steps taken per day before and after imputation##
```{r}
mean_diff <- impmean - normean
median_diff <- impmedian - normedian
```
*Mean difference is `r mean_diff`

*Median difference is `r median_diff`


##Difference of activity between weekday and weekend##
```{r}
Weekdays <- c("Monday","Tuesday","Wednesday","Thursday","Friday")

imputeddata$dayofweek <- 
  as.factor(ifelse(is.element(weekdays(as.Date(imputeddata$date)),Weekdays),"weekday","weekend"))
            
            
stepsbyintervalanddayofweek<-aggregate(steps ~ interval + dayofweek, imputeddata, mean)

xyplot(stepsbyintervalanddayofweek$steps ~ stepsbyintervalanddayofweek$interval|stepsbyintervalanddayofweek$dayofweek, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")


```